---
title: "Harnessing the power of {magick} in R for images' modifications"
author: "Aditya Dahiya"
date: "2024-11-25"
subtitle: "........................"
categories:
  - "#TidyTuesday"
  - "{magick}"
mage: "thumbnails/tidy_us_border_data.png"
format:
  html:
    code-fold: true
editor_options: 
  chunk_output_type: console
execute: 
  error: false
  message: false
  warning: false
  eval: false
filters:
  - social-share
share:
  permalink: "https://aditya-dahiya.github.io/projects_presentations/data_vizs.html"
  twitter: true
  linkedin: true
  email: true
  mastodon: true
bibliography: references.bib
comments:
  giscus: 
    repo: Aditya-Dahiya/projects_presentations
---

Harnessing the power of {magick} [@magick] for working with images in `R`.

```{r}
#| label: setup

# Loading the relevant packages
library(tidyverse)          # Data Wrangling
library(magick)             # For handling images
library(scales)             # Scales in the ggplot2
```

## Part 1: Transparent Overlays on images

The "Magick Overlay" [post](https://x.ai/grok/themockup.blog/posts/2022-06-22-magick-overlay/) on [The Mockup](https://themockup.blog/) blog by [Tom Mock](https://x.com/thomas_mock), dated June 22, 2022, explores the creation of dynamic, semi-transparent overlays using the `magick` package [@magick] in `R`, which interfaces with ImageMagick's powerful image processing capabilities. The technique involves **defining a custom overlay function, `create_overlay()`, which applies a coloured rectangle over an image with adjustable transparency**, enhancing the visual presentation for text overlays or focal points in images. This method leverages R's capabilities for image manipulation, offering flexibility beyond traditional CSS or HTML solutions by directly manipulating image layers. The process includes reading an image, applying color effects like `image_colorize()`, and compositing images with precise control over placement and transparency.

```{r}
# Load a temporary image to work with

url <- "https://upload.wikimedia.org/wikipedia/commons/f/ff/Tidyverse_hex_logo.png"
```
